Stochastic Approximation Monte Carlo with a Dynamic Update Factor (SAD)

### 2019-04-19 papers/histogram/sad.tex

The organization of this paper is as follows: In Section~\ref{sec:histogram}, we
describe in detail flat-histogram methods.  Section~\ref{sec:weight} outlines
weight-based histogram methods used in this work.

Section~\ref{sec:sad} introduces the dynamic stochastic approximation method.
In Section~\ref{sec:results}, we discuss the performance of the various methods
applied to the model system.

@@

The goal of flat histogram methods (also called \emph{broad histogram}
or \emph{multicanonical} methods) is to simulate each energy with
similar accuracy, so as to accurately determine the density of states
over a broad range of energies.  Once the density of states is known,
a number of thermodynamic quantities---such as heat capacity or
internal energy---can be easily computed for any temperature.
Properties that require more information---such as a spatial
correlation function or a response function---can still be computed
for any temperature, provided statistics are collected for each
individual energy, which can then be reweighted for any temperature.

Each flat histogram Monte Carlo method uses a given set of ``moves''
which change a system and should satisfy detailed balance.  Each
algorithm differs in how it determines the probability of accepting a
move, and in what additional statistics (if any) must be collected in
order decide on that probability.  The methods we compare belong to the
family of weight-based methods, which only require a set of
weights $w(E)$ be collected.

Weight-based methods calculate
the density of states $D(E)$ for discrete energy
levels~\cite{haber2018performance}. For this reason, energy binning
becomes an important consideration for complex continous systems.
Energy bins are typically of uniform size for the entire energy
continuum~\cite{fasnacht2004adaptive}. Some methods such as
AdaWL~\cite{koh2013dynamically}
employ a tunable mechanism for controlling the binning for low entropic
states in order to ensure the exploration of all energies.  The method
introduced in this paper is designed to scale appropriately as bin size
is changed, but we do not test this scaling, as we use a system with
discrete energy levels.

\section{Weight-based methods}\label{sec:weight}

We will begin by introducing three related weight-based methods which
rely on a weight function $w(E)$, noting that the method proposed in
this paper (discussed in Section~\ref{sec:sad}) also falls in this
category.  In these algorithms, the probability of accepting a move is
given by
\begin{equation}
	\mathcal{P}(E_\text{old} \rightarrow E_\text{new})
	= \min\left[1,\frac{w(E_\text{old})}{w(E_\text{new})}\right]
\end{equation}
the probability biases the simulation in favor of energies with low weights.
A set of weights that are proportional to the density of states of the
system $D(E)$ will result in an entirely flat histogram.  Thus
flat histogram is a criteria for convergence for these methods.  To avoid
overflow error , since the weights may vary over more than a
few hundred orders of magnitude, the natural logarithm of the weights are used
in this work.  Since $D(E) = S(E)$, the logarithm of the weights
can be thought of as an approximation of the entropy.

Each weighted density approach uses a random walk in energy space to
estimate the density of states.  The core of these approaches
is to continuously update the weights at each step of the simulation
\begin{equation}
	\ln{w_{t+1}(E)}=\ln{w_{t}(E)}
	+\gamma
\end{equation}
where $t$ is number of the current move, and $\gamma$ is an update
factor.  This update causes the random walk to avoid energies that
have been frequently sampled, leading to a rapid exploration of energy
space.  This approach, however, violates detailed balance, due to the
acceptance probabilities changing with each move.  The severity of
this violation decreased as we decrease $\gamma$.  The various
weight-based methods discussed here differ in how the decreasing of $\gamma$
is scheduled.

@@

\subsection{SAMC}
The Stochastic Approximation Monte Carlo (SAMC) algorithm addresses
the lack of convergence of Wang-Landau's approach with a simple
schedule by which the update factor $\gamma$ is continuously
decreased~\cite{liang2007stochastic, werlich2015stochastic,
  schneider2017convergence}.  The update factor is defined in the
original implementation~\cite{liang2007stochastic} in terms of an
arbitrary tunable parameter $t_0$.
\begin{align}
\gamma_{t}^{\text{SA}} =\frac{t_0}{\max(t_0,t)}
\end{align}
where as above $t$ is the number of moves that have been attempted.
The major advantage that SAMC offers is its proven convergence.
Provided the update factor satisfies
\begin{align}
\sum_{t=1}^\infty \gamma_{t} = \infty \quad\textrm{and}\quad
\sum_{t=1}^\infty \gamma_{t}^\zeta < \infty
\end{align}
where $\zeta \in \{1,2\}$, Liang has shown that the weights are proven
to converge to the true density of states~\cite{liang2006theory,
  liang2007stochastic}.  In addition, the energy range need not be
known a priori.  The time to converge depends only on the choice of
parameters $t_0$.  Unfortunately, there is no
prescription for finding an acceptable value for $t_0$, and
while the algorithm formally converges, for a poor choice of $t_0$
that convergence can be far too slow to be practical.
Liang et al. give a rule of thumb in which $t_0$
is chosen in the range from $2N_S$ to $100N_S$ where $N_S$ is the number
of energy bins~\cite{liang2007stochastic}.  Schneider et al. found
that for the
Ising model this heuristic is helpful for small spin
systems, but that larger systems require an even higher $t_0$
value~\cite{schneider2017convergence}.  Also, we find that for the square-well
fluid $t_0$ in some cases needs to be two orders of magnitude higher
than the rule of thumb of $100N_S$.

@@

\paragraph{Energy binning with SAD}

A significant advantage of SAD over other flat histogram methods is
that energy binning is solved dynamically. SAD should perform
independently for a reasonable range of binning because $\gamma \propto
\nicefrac{N_S}{t}$.  As the number of energy states found $N_S$
increases (fine binning), the time spent $t$ in each bin will decrease.
This argument explains why SAD should perform well for complex systems
in terms of binning. SAMC has a prefactor $\gamma_0$ to aid in a
similar way but this adds yet another parameter for the user to choose.

\paragraph{Scaling, step size, and SAD}
Methods such as WL can fail to converge for complex systems in part
because the same random MC moves are performed for the entire energy
range.  The main problem lies in trying to determine how to optimize
the step size such that the time to explore all energies is reasonable.
Small step sizes lead to improved sampling at the expense of slow
convergence. Koh et al. proposed multiplying $\gamma$ by an energy
dependent factor in order to improve the converge of WL in regards to
convergence and scaling~\cite{koh2013dynamically}.  Because the update
factor for SAD has an energy dependent factor already present (and
dynamically adjusted), we expect that SAD will perform consitently
scaling free for a reasonable choice of step size.  SAD should perform
extremely well compared with other flat histogram methods when applied
to complex continuous systems such as polymers/proteins and spin models.

\subsection{Convergence}
Because SAD scales as $1/t$ for large $t$, it satisfies the SAMC
convergence criteria.  However, we also want to ensure that SAD avoids
the exponentially poor convergence time that SAMC suffers when $t_0$ is
small as discussed in Section~\ref{sec:samc-convergence}.
To place a lower bound on the convergence of SAD, we will again
approximate the total entropy change using an integral.  To sidestep
the dynamical nature of SAD, we ensure that \emph{after discovering the
last important energy} at time $t_L$ the entropy change
$\Delta S_{\text{tot}}$ can be made in a reasonable number of iterations.
Thus we integrate the update factor from $t_L$ to some minimum time
$t_{\min}$ necessary for convergence.  Technically this is not a lower
bound as in the case of SAMC because the algorithm will already have a
reasonable approximation for the entropy.
\begin{align}
\Delta S_{\text{tot}} &= \int_{t_L}^{t_{\min}}
     \frac{
       3\langle S\rangle + \frac{t}{t_L}
     }{
       3\langle S\rangle + \frac{t}{N_S}\frac{t}{t_L}
     } dt \\
&\approx 3\Delta S_{\text{tot}} \left(1-\frac{t_L}{t_{\min}}\right)
\end{align}
where we assumed that $t_L\gg N_S\langle S\rangle$, which is to say
that it takes a long time to find all the interesting energy states.
Furthermore, we omitted the $\frac{t}{t_L}$ in the numerator, which is
of order unity.  This omission only makes our bound more conservative,
and enables us to solve to find that
\begin{align}
  t_{\min} &= \frac32 t_L
\end{align}
which means that we only need to run fifty percent longer after we
discover the final interesting energy state in order to have made
sufficient updates to \emph{possibly} converge to the true entropy.

@@

We use the average entropy error vs moves as a metric to compare
simulation runtimes and overall convergence. The overall accuracy
is determined by examining the fractional error of a particular method to
a precise reference system. For each simulation, the reference system
is chosen to be the final output of a SAMC simulation with a fixed interesting energy range.
We note this because SAMC does not require an energy range as an input parameter; however,
by fixing the energy range to for SAMC, a smaller $t_0$ can be chosen considerably reducing the
simulation time.
\begin{align}
\epsilon_\text{avg} = 1 - \bigg\lvert\frac{S_\text{method}}{S_\text{ref}}\bigg\rvert
\end{align}
Simulations that take more iterations to reach the same entropy error as a
different method take longer to converge. This behavior is seen in
Fig.~\ref{fig:fast-slow-gamma}.

@@

The methods SAD, WL, and $1/t$-WL compensate for the smaller
displacement distance as can be seen from Figure~\ref{fig:fast-slow-gamma}.
The update factors take approximately 10x longer to reach steady-state for
the slower translation scale. Because of this update behavior, these methods
do not suffer in converging to the correct density of states from having a
static parameter such as SAMC $t_0$ in the update factor. SAMC methods that converge
in a reasonable amount of moves in Figure~\ref{fig:fast-slow-gamma}a do not 
necessarily do so for Figure~\ref{fig:fast-slow-gamma}b.

Sad performs similarly regardless of the displacement distance chosen although
it transitions to $1/\sqrt{t}$ convergence after

\begin{figure}
\includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_50-ff0_17-N256-entropy-error-default.pdf}
  \caption{
  The average entropy error for each MC method for $N=256$,
               $\delta_0 = 0.05\sigma$, $\eta = 0.17$, and $T_{\min} = 1$
               as a function of number of iterations run.  The error is
               averaged over 8 independent simulations, and the best
               and worst simulations for each method are shown as a
               semi-transparent shaded area.  The update factor for this
               system is in Fig.~\ref{fig:gamma-vs-t} above.}\label{fig:n256}
\end{figure}
%\begin{figure}
%  \includegraphics[width=\columnwidth]{figs/N256-lndos-comparison}
%  \caption{\davidsays{This could be prettier, and do we need it? We
%      definitely need a caption if we keep it.} Use this to create
%    description of the system.}\label{fig:n256-lndos-comparison}
%\end{figure}

\subsection{A periodic system with 256 atoms}
For this simulation, we chose a minimum reduced
temperature of $1.0$, which corresponds to an interesting energy range
from $-915$ to $-509$.  The number of important energy states for this
system is then $N_S = E_{\min} - E_{\max} = 406$.
The entropy of this system is has an entropy that
differs by 275, corresponding to a ratio of
$10^{119}$ between the highest and lowest density of states.

Figure~\ref{fig:n256} shows the average error in the
entropy as a function of moves for this system with the reasonable
displacement distance of $\delta_0 = 0.05\sigma$.  The solid lines
represent the average of the absolute value of the error in the
entropy averaged over eight simulations using different random number
seeds.  The range of average errors for each simulation is shown as a
shaded region around its mean error.  By the time $10^8$ moves have
been made all but the SAMC simulation with the shortest $t_0$ have
begun to converge as $1/\sqrt{t}$.  We then see the WL error begin to
saturate around $10^{10}$ moves.

For the larger system, SAMC parameter values of $t_0$ that gave rapid
convergence for the $N = 50$ system with $\delta = 0.05\sigma$ do not
converge rapidly.  Larger $t_0$ parameters are simulated for
comparison with SAD.  SAD has significantly smaller fluctuations
in the average entropy error for a given number of moves.

@@

\subsubsection{Discussion on convergence and update factor}

A close examination of Fig.~\ref{fig:gamma-vs-t} shows that $\gamma$
for the Wang-Landau algorithm decreases much more rapidly than
$\nicefrac{1}{\sqrt{t}}$. It can be observed  for the simulations that
were carried out that all the methods from the weight-based family
struggle to converge if the update factor decreases too rapidlly.
Fig.~\ref{fig:n256} shows SAMC with inappropriate choices
of $t_0$ i.e. less than $\nicefrac{10^4}{\sqrt{t}}$ failing to converge
in a reasonable amount of time. Thus, the update factor $\gamma$, shown
as a function of simulation moves in Fig.~\ref{fig:gamma-vs-t}, proves
useful in determining which methods will converge in a reasonable
amount of time. We compare each method to a reference SAMC
simulation for an ideal choice of $t_0$.  This gives an unbiased
comparison with our developed method SAD. The dynamic version of SAMC
which we call SAD performs as well as SAMC (assuming the user can
correctly identify an appropriate choice of $t_0$ for SAMC).  SAD
outperforms WL and 1/t-WL? for the same given number of moves. The
absence of user defined parameters makes SAD an exciting alternative to
methods where user experience or successive implementation iteration
are needed to achieve reasonable results.

@@

>>>>

The organization of this paper is as follows: In Section~\ref{sec:histogram}, we
describe in detail flat-histogram methods.
%
Section~\ref{sec:sad} introduces the dynamic stochastic approximation method.
In Section~\ref{sec:results}, we discuss the performance of the various methods
applied to the model system.

@@

The goal of flat histogram methods (also called \emph{broad histogram}
or \emph{multicanonical} methods) is to simulate each energy with
similar accuracy, so as to accurately determine the density of states
over a broad range of energies---and thus to determine the thermodynamic
quantities such as heat capacity or internal energy
over a broad range of temperatures.
Properties that require more information---such as a spatial
correlation function or a response function---can still be computed
for any temperature, provided statistics are collected for each
individual energy, which can then be reweighted for any temperature.

Each flat histogram Monte Carlo method uses a given set of ``moves''
which change a system and should satisfy detailed balance.  Each
algorithm differs in how it determines the probability of accepting a
move, and in what additional statistics must be collected in
order decide on that probability.

Flat histogram methods calculate
the density of states $D(E)$ for a discrete set of
energies~\cite{haber2018performance}. For this reason, energy binning
becomes an important consideration for systems with a continuum of
energies possible.
Energy bins are typically of uniform size for the entire energy
continuum~\cite{fasnacht2004adaptive}. Some methods such as
AdaWL~\cite{koh2013dynamically}
employ a tunable mechanism for controlling the binning for low entropic
states in order to ensure the exploration of all energies.  The method
introduced in this paper is designed to scale appropriately as bin size
is changed, but we do not test this scaling, as we use a system with
discrete energy levels.

In this section we will introduce four related flat histogram methods
which each rely on a weight function $w(E)$, noting that the method proposed in
this paper (discussed in Section~\ref{sec:sad}) also falls in this
category.  In these algorithms, the probability of accepting a move is
given by
\begin{equation}
	\mathcal{P}(E_\text{old} \rightarrow E_\text{new})
	= \min\left[1,\frac{w(E_\text{old})}{w(E_\text{new})}\right]
\end{equation}
the probability biases the simulation in favor of energies with low weights.
A set of weights that are proportional to the density of states of the
system $D(E)$ will result in an entirely flat histogram.  Thus
flat histogram is a criteria for convergence for these methods.  To avoid
overflow error , since the weights may vary over more than a
few hundred orders of magnitude, the natural logarithm of the weights are used
in this work.  Since $D(E) = S(E)$, the logarithm of the weights
can be thought of as an approximation of the entropy.

Each approach uses a random walk in energy space to
estimate the density of states.  The core of these approaches
is to continuously update the weights at each step of the simulation
\begin{equation}
	\ln{w_{t+1}(E)}=\ln{w_{t}(E)}
	+\gamma
\end{equation}
where $t$ is number of the current move, and $\gamma$ is an update
factor.  This update causes the random walk to avoid energies that
have been frequently sampled, leading to a rapid exploration of energy
space.  This approach, however, violates detailed balance, due to the
acceptance probabilities changing with each move.  The severity of
this violation decreased as we decrease $\gamma$.  The various
weight-based methods discussed here differ in how the decreasing of $\gamma$
is scheduled.

@@

\subsection{SAMC}
The Stochastic Approximation Monte Carlo (SAMC) algorithm addresses
the lack of convergence of Wang-Landau's approach with a simple
schedule by which the update factor $\gamma$ is continuously
decreased~\cite{liang2007stochastic, werlich2015stochastic,
  schneider2017convergence}.  The update factor is defined in the
original implementation~\cite{liang2007stochastic} in terms of an
arbitrary tunable parameter $t_0$.
\begin{align}
\gamma_{t}^{\text{SA}} =\frac{t_0}{\max(t_0,t)}
\end{align}
where as above $t$ is the number of moves that have been attempted.
The major advantage that SAMC offers is its proven convergence.
Provided the update factor satisfies
\begin{align}
\sum_{t=1}^\infty \gamma_{t} = \infty \quad\textrm{and}\quad
\sum_{t=1}^\infty \gamma_{t}^\zeta < \infty
\end{align}
where $\zeta \in \{1,2\}$, Liang has shown that the weights are proven
to converge to the true density of states~\cite{liang2006theory,
  liang2007stochastic}.  In addition, the energy range need not be
known a priori.  The time to converge depends only on the choice of
parameter $t_0$.  Unfortunately, there is no
prescription for finding an acceptable value for $t_0$, and
while the algorithm formally converges, for a poor choice of $t_0$
that convergence can be far too slow to be practical.
Liang et al. give a rule of thumb in which $t_0$
is chosen in the range from $2N_S$ to $100N_S$ where $N_S$ is the number
of energy bins~\cite{liang2007stochastic}.  Schneider et al. found
that for the
Ising model this heuristic is helpful for small spin
systems, but that larger systems require an even higher $t_0$
value~\cite{schneider2017convergence}.  Also, we find that for the square-well
fluid $t_0$ in some cases needs to be two orders of magnitude higher
than the rule of thumb of $100N_S$.

@@

%~ \paragraph{Energy binning with SAD}

A significant advantage of SAD over other flat histogram methods is
that energy binning is handled dynamically. SAD should perform
independently for a reasonable range of binning because $\gamma \propto
\nicefrac{N_S}{t}$.  As the number of energy states found $N_S$
increases (fine binning), the time spent $t$ in each bin will decrease
with the effect that the convergence should be roughly independent of
the bin size chosen.
This argument explains why SAD should perform well for continuum systems
in terms of binning. SAMC has a prefactor $\gamma_0$ to aid in a
similar way but this adds yet another parameter for the user to choose.

%~ \paragraph{Scaling, step size, and SAD}
%~ Methods such as WL can fail to converge for complex systems in part
%~ because the same random MC moves are performed for the entire energy
%~ range.  The main problem lies in trying to determine how to optimize
%~ the step size such that the time to explore all energies is reasonable.
%~ Small step sizes lead to improved sampling at the expense of slow
%~ convergence. Koh et al. proposed multiplying $\gamma$ by an energy
%~ dependent factor in order to improve the converge of WL in regards to
%~ convergence and scaling~\cite{koh2013dynamically}.  Because the update
%~ factor for SAD has an energy dependent factor already present (and
%~ dynamically adjusted), we expect that SAD will perform consitently
%~ scaling free for a reasonable choice of step size.  SAD should perform
%~ extremely well compared with other flat histogram methods when applied
%~ to complex continuous systems such as polymers/proteins and spin models.

%~ \subsection{Convergence}
%~ Because SAD scales as $1/t$ for large $t$, it satisfies the SAMC
%~ convergence criteria.  However, we also want to ensure that SAD avoids
%~ the exponentially poor convergence time that SAMC suffers when $t_0$ is
%~ small as discussed in Section~\ref{sec:samc-convergence}.
%~ To place a lower bound on the convergence of SAD, we will again
%~ approximate the total entropy change using an integral.  To sidestep
%~ the dynamical nature of SAD, we ensure that \emph{after discovering the
%~ last important energy} at time $t_L$ the entropy change
%~ $\Delta S_{\text{tot}}$ can be made in a reasonable number of iterations.
%~ Thus we integrate the update factor from $t_L$ to some minimum time
%~ $t_{\min}$ necessary for convergence.  Technically this is not a lower
%~ bound as in the case of SAMC because the algorithm will already have a
%~ reasonable approximation for the entropy.
%~ \begin{align}
%~ \Delta S_{\text{tot}} &= \int_{t_L}^{t_{\min}}
     %~ \frac{
       %~ 3\langle S\rangle + \frac{t}{t_L}
     %~ }{
       %~ 3\langle S\rangle + \frac{t}{N_S}\frac{t}{t_L}
     %~ } dt \\
%~ &\approx 3\Delta S_{\text{tot}} \left(1-\frac{t_L}{t_{\min}}\right)
%~ \end{align}
%~ where we assumed that $t_L\gg N_S\langle S\rangle$, which is to say
%~ that it takes a long time to find all the interesting energy states.
%~ Furthermore, we omitted the $\frac{t}{t_L}$ in the numerator, which is
%~ of order unity.  This omission only makes our bound more conservative,
%~ and enables us to solve to find that
%~ \begin{align}
  %~ t_{\min} &= \frac32 t_L
%~ \end{align}
%~ which means that we only need to run fifty percent longer after we
%~ discover the final interesting energy state in order to have made
%~ sufficient updates to \emph{possibly} converge to the true entropy.

@@

We use the average entropy error vs moves as a metric to compare
simulation runtimes and overall convergence. The overall accuracy
is determined by examining the fractional error of a particular method to
a precise reference system. For each simulation, the reference system
is chosen to be the final output of a SAMC simulation with a fixed energy range
corresponding to the temperature range of interest.
Although SAMC does not require an energy range as an input parameter,
we find that by limiting the simulation to this energy range, we can
acheive much faster convergence with a smaller $t_0$.  We compute an average
of the error by averaging the error in the entropy over the 
interesting energy range, and then averaging \emph{this} error over several
simulations run with different random number seeds.
%~ \begin{align}
%~ \epsilon_\text{avg} =
  %~ \frac{1}{N_S \text{num seeds}}\sum_s^{\text{num seeds}}\sum_{i}^{N_S} \left|S_\text{ref}(E_i) - S_\text{method}(E_i)\right|
%~ \end{align}

@@

The methods SAD, WL, and $1/t$-WL compensate for the smaller
displacement distance as can be seen from Figure~\ref{fig:fast-slow-gamma}.
The update factors take approximately 10x longer to reach steady-state for
the slower translation scale. Because of this update behavior, these methods
are able to converge regardless of the displacement distance used in Monte
Carlo moves.

%~ SAMC methods that converge
%~ in a reasonable amount of moves in Figure~\ref{fig:fast-slow-gamma}a do not 
%~ necessarily do so for Figure~\ref{fig:fast-slow-gamma}b.
%
%~ Sad performs similarly regardless of the displacement distance chosen although
%~ it transitions to $1/\sqrt{t}$ convergence after

\begin{figure}
\includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_50-ff0_17-N256-entropy-error-default.pdf}
  \caption{
  The average entropy error for each MC method for $N=256$,
               $\delta_0 = 0.05\sigma$, $\eta = 0.17$, and $T_{\min} = 1$
               as a function of number of iterations run.  The error is
               averaged over 8 independent simulations, and the best
               and worst simulations for each method are shown as a
               semi-transparent shaded area.  The update factor for this
               system is in Fig.~\ref{fig:gamma-vs-t} above.}\label{fig:n256}
\end{figure}
%~ \begin{figure}
  %~ \includegraphics[width=\columnwidth]{figs/N256-lndos-comparison}
  %~ \caption{\davidsays{This could be prettier, and do we need it? We
      %~ definitely need a caption if we keep it.} Use this to create
    %~ description of the system.}\label{fig:n256-lndos-comparison}
%~ \end{figure}

\subsection{A periodic system with 256 atoms}
Next we will introduce a considerably larger simulation containing
256 atoms which has a total entropy range of around 1500.  This makes
exploring the entire range of energies extremely expensive, and strongly
favors methods that restrict the energy (or temperature) range of
interest.
For this simulation, we chose a much higher minimum reduced
temperature of $1.0$, which corresponds to an interesting energy range
from $-915$ to $-509$.  The number of important energy states for this
system is then $N_S = E_{\min} - E_{\max} + 1 = 407$.
The entropy of this system is has an entropy that
differs by 275, corresponding to a ratio of
$10^{119}$ between the highest and lowest density of states.
\davidsays{Also give the entropy difference, looking at the data.}
\jpsays{$395.2833013702813 - 4.754731035740479e^{-8}$ = $395.2833$ so $2^{396.2833}$
gives us about $10^{118}$ from 'samc-1e4-256-s8.yaml'}

Figure~\ref{fig:n256} shows the average error in the
entropy as a function of moves for this system with the reasonable
displacement distance of $\delta_0 = 0.05\sigma$.  The solid lines
represent the average of the absolute value of the error in the
entropy averaged over eight simulations using different random number
seeds.  The range of average errors for each simulation is shown as a
shaded region around its mean error.  By the time $10^8$ moves have
been made all but the SAMC simulation with the shortest $t_0$ have
begun to converge as $1/\sqrt{t}$.  We then see the WL error
saturate around $10^{10}$ moves.

%~ For this larger system, SAMC parameter values of $t_0$ that gave rapid
%~ convergence for the $N = 50$ system with $\delta = 0.05\sigma$ do not
%~ converge rapidly.

@@

%~ \subsection{Discussion on convergence and update factor}

%~ A close examination of Fig.~\ref{fig:gamma-vs-t} shows that $\gamma$
%~ for the Wang-Landau algorithm decreases much more rapidly than
%~ $\nicefrac{1}{t}$. It can be observed  for the simulations that
%~ were carried out that all the methods
%~ struggle to converge if the update factor decreases too rapidlly.
%~ Fig.~\ref{fig:n256} shows SAMC with inappropriate choices
%~ of $t_0$ i.e. less than $\nicefrac{10^4}{\sqrt{t}}$ failing to converge
%~ in a reasonable amount of time. Thus, the update factor $\gamma$, shown
%~ as a function of simulation moves in Fig.~\ref{fig:gamma-vs-t}, proves
%~ useful in determining which methods will converge in a reasonable
%~ amount of time. We compare each method to a reference SAMC
%~ simulation for an ideal choice of $t_0$.  This gives an unbiased
%~ comparison with our developed method SAD. The dynamic version of SAMC
%~ which we call SAD performs as well as SAMC (assuming the user can
%~ correctly identify an appropriate choice of $t_0$ for SAMC).  SAD
%~ outperforms WL and 1/t-WL? for the same given number of moves. The
%~ absence of user defined parameters makes SAD an exciting alternative to
%~ methods where user experience or successive implementation iteration
%~ are needed to achieve reasonable results.

@@

\davidsays{We find both 1/tWL and sad do really well and very similarly.
SAD wins if you want to specify a temperature range rather than an energy.
1/tWL cheated because you had to run a prior simulation in order to find
the energy range of interest.  Yay good stuff.}

\jpsays{Both SAD and 1/t-WL converge more rapidly than Wl or SAMC. SAD requires
the user to specify a temperature range rather than an energy range as 1/t-WL
does. In addition to requiring the energy range,
1/t-WL requires a prior simulation in order to determine the energy range
of interest. WL saturates in convergence as discussion in the introduction and
requires the user to optimize the choice of parameters to avoid this such as
increasing the flatness criterion.  SAMC converges for a reasonable choice of
$t_0$ but this parameter can be difficult to tune especially across different
energy bins, displacement distances, or systems. Additionaly, SAMC does not
converge as rapidly as either SAD or 1/t-WL even for the best choice of $t_0$.}
